---
title: 'Practical Machine Learning: Course Project'
output: html_document
---

Pin Choengtawee

June 19, 2016

#Exceutive Summary

The goal of this report was building a predictive model to predict the manner in which participants did the exercise by using data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. After performed data loading and data cleaning, first step was to devide data into two sets: which are trainging set and testing set. Next was to build a fit model by using the Random Forest method and apply the select model to the test set. As a result, The predictive model developed using Random Forest was able to achieve over 99.99% accuracy, or less than 0.03% out-of-sample error, and was able to predict the 20 test cases with 100% accuracy.

#Data Loading  and Exploratory Analysis

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r}
df <- "pml_training.csv"
if (!file.exists(df)) {
    url <-
        "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
    df <- "pml_training.csv"
    download.file(url, destfile = df)
}
training <- read.csv(df, na.strings = c("NA","#DIV/0!",""))

df <- "pml_testing.csv"
if (!file.exists(df)) {
    url <-
        "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
    download.file(url, destfile = df)
}
testing <- read.csv(df, na.strings = c("NA","#DIV/0!",""))
```

#Data Cleaning

Data cleaning includes removing columns with Near Zero Values, Removing columns with NA or is empty, and removing V1 which seems to be a serial number.

```{r}
library(caret)

# Remove columns with Near Zero Values
subTrain <-
    training[, names(training)[!(nzv(training, saveMetrics = T)[, 4])]]

# Remove columns with NA or is empty
subTrain <-
    subTrain[, names(subTrain)[sapply(subTrain, function (x)
        ! (any(is.na(x) | x == "")))]]

# Remove V1 which seems to be a serial number
subTrain <- subTrain[,-1]
subTrain <- subTrain[, c(1:3, 5:58)]
```

#Prediction Model Building

Befor building the prediction model, devide the data set into two sets: a training set and a validation/test set

```{r}
inTrain <- createDataPartition(subTrain$classe, p = 0.6, list = FALSE)
TrainSet <- subTrain[inTrain,]
ValidationSet <- subTrain[-inTrain,]
dim(TrainSet)
dim(ValidationSet)

```
      
###a) Building a fit model by using a Random Forest method

```{r}
#install.packages("doParallel")
library(doParallel)
# Check if model file exists
model <- "modelFit.RData"
if (!file.exists(model)) {
    
    # If not, set up the parallel clusters.  
    require(parallel)
    require(doParallel)
    cl <- makeCluster(detectCores() - 1)
    registerDoParallel(cl)
    
    fit <- train(TrainSet$classe ~ ., method = "rf", data = TrainSet)
    save(fit, file = "modelFit.RData")
    
    stopCluster(cl)
} else {
    # Good model exists from previous run, load it and use it.  
    load(file = "modelFit.RData", verbose = TRUE)
}
```

###b) Applying the selected model to the validation/test data

```{r}
predTrain <- predict(fit, TrainSet)
confusionMatrix(predTrain, TrainSet$classe)
predValidation <- predict(fit, ValidationSet)
confusionMatrix(predValidation, ValidationSet$classe)
```

From the validation subset, the accuracy is still very high, at above 99%. Given the level of accuracy, there is no need to build another prediction model for better accuracy or to stack multiple prediction models. The following is the lists of important predictors in the model.

```{r}
varImp(fit)
```

```{r}
fit$finalModel
```

Even though the reported OOB Estimated Error is at 13%, the validation accuracy is high at over 99% and Cross-Validation out-of-sample error rate is only 0.03%. Thus, the prediction model should be applied to the final testing set and predict the classe in the 20 test cases.

```{r}
predTesting <- predict(fit, testing)
predTesting
```


#Create Files 

```{r}
pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
}

pml_write_files(predTesting)
```